### 2021-05-10 writeups/Su_weak_tides.tex
stash 0512

% vim 0_eta/1*.py 0_eta/3*.py 0_eta/6*.py 3_toy3/1sim.py 99_misc/0*.py 99_misc/2*.py 99_misc/5*.py 1_weaktide/5*.py 1_weaktide/6*.py
% cd 0_eta && python 1*.py && python 3*.py && python 6*.py && cd .. && cd 3_toy3 && python 1sim.py && cd .. && cd 99_misc && python 0*.py && python 2*.py && python 5*.py && cd .. && cd 1_weaktide && python 5*.py && python 6*.py
    \documentclass[
        fleqn,
        usenatbib,
        % referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        ae, aecompl,
        graphicx,
        booktabs,
        xcolor,
    }

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2#1}{\partial#2^2}}
    % inline
    \newcommand*{\mdil}[2]{\mathrm{D}#1/\mathrm{D}#2}
    \newcommand*{\pdil}[2]{\partial#1/\partial#2}
    \newcommand*{\rdil}[2]{\mathrm{d}#1/\mathrm{d}#2}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\left\langle#1\right\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \newcommand*{\bm}[1]{\mathbf{#1}}
    \newcommand*{\uv}[1]{\hat{\mathbf{#1}}}
    \newcommand*{\md}[0]{\mathrm{d}}
    \DeclareMathOperator*{\med}{med}
    \DeclareMathOperator*{\erf}{erf}

\title[Weak Tides and Cassini States]{Dynamics of Colombo's Top: Tidal
Dissipation and Resonance Capture}
\author[Y. Su and D. Lai.]{
Yubo Su,$^1$\thanks{E-mail: yubosu@astro.cornell.edu},
Dong Lai$^{1,2}$
\\
$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
Astronomy, Cornell University, Ithaca, NY 14853, USA\\
$^2$ Tsung-Dao Lee Institute \& School of Physics and Astronomy, Shanghai Jiao
Tong University, 200240 Shanghai, China
}

\date{Accepted XXX\@. Received YYY\@; in original form ZZZ}

\pubyear{2021}

\begin{document}\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

\begin{abstract}
    Abstract here
\end{abstract}

\begin{keywords}
planet-star interactions
\end{keywords}

\section{Introduction}\label{s:intro}

% (i) What is obliquity? Why is it important, esp in exoplanetary systems?
% (ii) what are the dynamics of obliquity?
% (iii) Introduce Paper I

It is well recognized that the obliquity of a planet, the angle between the spin
and orbital axes, likely reflects its dynamical history. In our Solar
System, planetary obliquities (hereafter just ``obliquities'') range from
$3.1^\circ$ for Jupiter to $26.7^\circ$ for Saturn to $98^\circ$ for Uranus. The
obliquities of exoplanets are challinging to measure, and so far only loose
constraints have been obtained for the obliquity of a faraway ($\gtrsim
50\;\mathrm{AU}$) planetary-mass companion \citep{bryan2020obliquity}.
Nevertheless, there are prospects for better constraints on exoplanetary
obliquities in the coming years, such as using high-resolution spectroscopy to
measure $v\sin i$ for a planet \citep{snellen2014fast, bryan2018constraints} and
using high-precision photometry to measure the asphericity of a planet
\citep{seager2002constraining}. Substantial obliquities are of increasing
theoretical interest for their proposed role in explaining peculiar thermal
phase curves \citep[see e.g.][]{millholland_signatures, ohno_infer_obl}, in
enhancing tidal dissipation in hot Jupiters \citep{millholland_wasp12b} and
super-Earths \citep{millholland2019obliquity}, and in the formation of
ultra-short-period planets \citep[USPs;][]{millholland2020formation} .

While nonzero obliquities are sometimes attributed to one or many giant impacts
\citep[e.g.]{original_gi, benz1989tilting, korycansky1990one, morbidelli_gi},
some studies suggest that large planetary obliquities may be due to encountering
spin-orbit resonances. In this scenario, a rotating planet is subjected to a
gravitational torque from its host star, making its spin axis precess around its
orbital (angular momentum) axis. At the same time, the orbital axis precesses
around another fixed axis under the gravitational influence of other masses in
the system, e.g.\ additional planets or a protoplanetary disk. When the two
precession frequencies become commensurate, a resonance can occur that excites
the obliquity to large values. This model is known as ``Colombo's Top'' after
the seminal work of \citet{colombo1966}, and subsequent works have investigated
the rich dynamics of this system \citep{peale1969, peale1974possible,
ward1975tidal, henrard1987}. Such resonances have been invoked to explain the
obliquities of both the Solar System gas giants \citep{ward2004I, ward2004II,
ward_jupiter, vokrouhlicky2015tilting, saillenfest2020future,
saillenfest2021large} and the ice giants \citep{hamilton_tilting_ice}.

In a previous paper \citep[hereafter Paper I]{su2020}, we presented a systematic
and general investigation of the dynamics of Colombo's Top when the two
precession frequencies of the system evolve through a commensurability. We
applied our results to investigate the generation of exoplanetary obliquities
via a dissipating protoplanetary disk. However, our model did not consider the
effect of additional torques in the system. Often, tidal dissipation in the
planet both causes the planet's spin frequency to approach its orbital frequency
and drives the planet's spin axis towards its orbital axis, complicating the
evolution of Colombo's Top \citep{fabrycky_otides, peale2008obliquity}. In this
paper, we extend the approach of Paper I to include tidal dissipation. Our new
results include a simplified formulation of the effect of tidal dissipation on
the equilibria of the system (Cassini States) and a novel, complete description
of the resonance encounter process.

In Section~\ref{s:theory}, we briefly review Colombo's Top. In
Section~\ref{s:toy_model}, we investigate the effect of adding a simple
alignment torque to Colombo's Top. The resulting dynamics are illustrative of
the new behavior that emerges due to tidal dissipation. In
Section~\ref{s:full_tide_prob}, we solve for the dynamics of the system
including a realistic model of tidal dissipation. In Section~\ref{s:summary}, we
apply our results to three astrophysical scenarios of interest: (i) a super
Earth with a cold Jupiter companion, (ii) the formation of USPs, and (iii) the
rapid orbital decay of WASP-12b. We summarize and discuss in
Section~\ref{s:summary}.

\section{Spin Evolution Equations and Cassini States: Review}\label{s:theory}

In this section, we briefly introduce the spin dynamics of the planet (for a
more detailed treatment, see Paper I). We consider a star of mass $M_\star$
hosting an inner oblate planet of mass $m$ and radius $R$ on a circular orbit
with semi-major axis $a$ and an outer perturber of mass $m_{\rm p}$ on a
circular orbit with semi-major axis $a_{\rm p}$. We assume that the two orbits
are mutually inclined by the angle $I$. Denote $\bm{S}$ the spin angular
momentum and $\bm{L}$ the orbital angular momentum of the planet, and
$\bm{L}_{\rm p}$ the angular momentum of the perturber. The corresponding unit
vectors are $\uv{s} \equiv \bm{S} / S$, $\uv{l} \equiv \bm{L} / L$, and
$\uv{l}_{\rm p} \equiv \bm{L}_{\rm p} / L_{\rm p}$. The spin axis $\uv{s}$ of
the planet tends to precess around its orbital (angular momentum) axis $\uv{l}$,
driven by the gravitational torque from the host star acting on the planet's
rotational bulge. On the other hand, $\uv{l}$ and the disk axis $\uv{l}_{\rm p}$
precess around each other due to gravitational interactions. We assume $S \ll L
\ll L_{\rm p}$, so $\uv{l}_{\rm p}$ and $\uv{l}$ are nearly constant. The
equations of motion for $\uv{s}$ and $\uv{l}$ in this limit are
(\citealp{anderson2018teeter}, Paper I)
\begin{align}
    \rd{\uv{s}}{t}
        &= \omega_{\rm sl}\p{\uv{s} \cdot \uv{l}}\p{\uv{s} \times \uv{l}}
        \equiv \alpha\p{\uv{s} \cdot \uv{l}}\p{\uv{s} \times
        \uv{l}},\label{eq:dsdt1}\\
    \rd{\uv{l}}{t} &= \omega_{\rm lp}\p{\uv{l} \cdot \uv{l}_{\rm p}}\p{\uv{l}
        \times \uv{l}_{\rm p}} \equiv -g\p{\uv{l} \times \uv{l}_{\rm p}},
        \label{eq:dldt1}
\end{align}
where
\begin{align}
    \omega_{\rm sl} &\equiv \frac{3GJ_2 mR^2 M_\star}{2a^3 I\Omega_{\rm s}}
        = \frac{3k_q}{2k}\frac{M_\star}{m}\p{\frac{R}{a}}^3 \Omega_{\rm s},
            \label{eq:wsl}\\
    \omega_{\rm lp} &= \frac{3m_{\rm p}}{4M_\star}\p{\frac{a}{a_{\rm p}}}^3
        n.\label{eq:wlp}
\end{align}
In Eq.~\eqref{eq:wsl}, $\Omega_{\rm s}$ is the spin frequency of the inner
planet, $\mathcal{I} = k mR^2$ (with $k$ a constant) is its moment of inertia
and $J_2 = k_{\rm q}\Omega_{\rm s}^2 (R^3/Gm)$ (with $k_{q}$ a constant) is its
rotation-induced (dimensionless) quadrupole moment [for a body with uniform
density, $k=0.4, k_{\rm q}=0.5$; for rocky planets, $k\simeq 0.2$ and $k_{\rm
q}\simeq 0.17$ \citep[e.g.][]{lainey2016quantification}]. In other studies,
$3k_{\rm q} / 2 k$ is often notated as $k_2 / 2C$
\citep[e.g.][]{millholland_disk}. In Eq.~\eqref{eq:wlp}, $n \equiv
\sqrt{GM_\star/a^3}$ is the inner planet's orbital mean motion,  and we have
assumed $a_{\rm p}\gg a$ and included only the leading-order (quadrupole)
interaction between the inner planet and perturber. Following standard notation,
we define $\alpha = \omega_{\rm sl}$ and $g \equiv -\omega_{\rm 1p} \cos I$
\citep[e.g.][]{colombo1966}.

As in Paper I, we combine Eqs.~(\ref{eq:dsdt1}--\ref{eq:dldt1}) into a
single equation by transforming into a frame rotating about $\uv{l}_{\rm p}$
with frequency $g$. In this frame, $\uv{l}_{\rm p}$ and $\uv{l}$ are both fixed,
and $\uv{s}$ evolves as:
\begin{equation}
    \p{\rd{\uv{s}}{t}}_{\rm rot}
        = \alpha\p{\uv{s} \cdot \uv{l}}\p{\uv{s} \times \uv{l}}
            + g\p{\uv{s} \times \uv{l}_{\rm p}}. \label{eq:dsdt_rot}
\end{equation}
In this reference frame, we choose the coordinate system such that $\uv{z} =
\uv{l}$ and $\uv{l}_{\rm p}$ lies in the $\uv{x}$-$\uv{z}$ plane. We describe
$\uv{s}$ in spherical coordinates using the polar angle $\theta$, the planet's
obliquity, and $\phi$, the precessional phase of $\uv{s}$ about $\uv{l}$,
defined so that when $\phi = 0^\circ$, $\uv{l}_{\rm p}$ and $\uv{s}$ are on
opposide sides of $\uv{l}$.

The equilibria of Eq.~\eqref{eq:dsdt_rot} are referred to as \emph{Cassini
States} \citep[CSs;][]{colombo1966, peale1969}. We follow the notation of
Paper I, where the parameter
\begin{equation}
    \eta \equiv -\frac{g}{\alpha},\label{eq:def_eta}
\end{equation}
is introduced. For a given value of $\eta$, there can be either two or four CSs,
all of which require $\uv{s}$ lie in the plane of $\uv{l}$ and $\uv{l}_{\rm p}$.
Following the standard nomenclature, $\uv{s}$ and $\uv{l}_{\rm p}$ are on
opposite sides of $\uv{l}$ for CSs 1, 3, and 4, and are on the same side for
CS2. We depart from the standard convention and simply label the CSs in
spherical coordinates: Figure~\ref{fig:cs_locs} shows the CS obliquities as a
function of $\eta$. CS1 and CS4 do not exist when $\eta > \eta_{\rm c}$, where
\begin{equation}
    \eta_{\rm c} \equiv \p{\sin^{2/3}\!I + \cos^{2/3}\!I}^{-3/2}.
        \label{eq:def_etac}
\end{equation}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/2_cs_locs_phi.png}
    \caption{Cassini State obliquities $\theta$ as a function of $\eta \equiv
    -g/\alpha$ (Eq.~\ref{eq:def_eta}) for $I = 20^\circ$. The vertical dashed
    line denotes $\eta_{\rm c}$, where the number of Cassini States changes from
    four to just two (Eq.~\ref{eq:def_etac}). The y-axis labels on the right of
    the plot show the asymptotic obliquities for CS2 and CS3, $I$ and $180^\circ
    - I$ respectively. Note that $\theta$ does not follow the standard
    convention \citep[e.g.][]{colombo1966} and is simply the angle between
    $\uv{s}$ and $\uv{l}$.}\label{fig:cs_locs}
\end{figure}

The Hamiltonian corresponding to Eq.~\eqref{eq:dsdt_rot} is
\begin{align}
    H &= -\frac{\alpha}{2}\p{\uv{s} \cdot \uv{l}}^2
            - g\p{\uv{s} \cdot \uv{l}_{\rm d}}\nonumber\\
        &= -\frac{\alpha}{2} \cos^2\theta
            - g\p{\cos\theta \cos I - \sin I \sin\theta \cos \phi}.\label{eq:H}
\end{align}
Here, $\cos \theta$ and $\phi$ form a canonically conjugate pair of variables.
Figure~\ref{fig:1contours} shows the level curves of this Hamiltonian for $I =
20^\circ$, for which $\eta_{\rm c} \approx 0.574$ (Eq.~\ref{eq:def_etac}). When $\eta
< \eta_{\rm c}$, CS4 exists and is a saddle point. The infinite-period orbits
originating and ending at CS4 form the \emph{separatrix} and divide phase space
into three zones. $\phi$ librates for trajectories in zone II and circulates for
trajectories in zones I and III\@. On the other hand, when $\eta > \eta_{\rm
c}$, the separatrix is absent and all trajectories circulate. When the
separatrix exists, we divide it into two curves: $\mathcal{C}_+$, the boundary
between zones I and II, and $\mathcal{C}_-$, the boundary between zones II and
III\@.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/0_eta/1contours20.png}
    \caption{Level curves of the Cassini State Hamiltonian (Eq.~\ref{eq:H}) for
    $I = 20^\circ$, for which $\eta_{\rm c} \approx 0.57$
    (Eq.~\ref{eq:def_etac}). For $\eta < \eta_{\rm c}$, there are four Cassini
    States (labeled), while for $\eta > \eta_{\rm c}$ there are only two. In the
    former case, the existence of a \emph{separatrix} (solid black lines)
    separates phase space into three numbered zones (I/II/III, labeled). We
    denote the upper and lower legs of the separatrix by $\mathcal{C}_{\pm}$
    respectively, as shown in the upper two panels. }\label{fig:1contours}
\end{figure}

\section{Spin Evolution with Alignment Torque}\label{s:toy_model}

In this section, we consider a simplified dissipative torque that isolates the
important new phenomenon presented in this paper. We assume that the spin
magnitude of the planet is constant, so $\alpha$ and $g$ are both fixed, while
the spin orientation $\uv{s}$ experiences an alignment torque towards $\uv{l}$
on the alignment timescale $t_{\rm al}$:
\begin{equation}
    \p{\rd{\uv{s}}{t}}_{\rm tide}
        = \frac{1}{t_{\rm al}} \uv{s} \times \p{\uv{l} \times \uv{s}}.
        \label{eq:dsdt_tide_toy}
\end{equation}
The full equations of motion for $\uv{s}$ in the coordinates $\theta$ and $\phi$
can be written:
\begin{align}
    \rd{\theta}{t} &= -g\sin I \sin \phi - \frac{1}{t_{\rm al}} \sin \theta,
        \label{eq:dqdt_toy}\\
    \rd{\phi}{t} &= -\alpha \cos\theta
        - g\p{\cos I + \sin I \cot \theta \cos \phi}.\label{eq:dfdt_toy}
\end{align}

\subsection{Modified Cassini States and Linear Stability
Analysis}\label{ss:tidal_equils}

If the alignment torque is weak ($\abs{g}t_{\rm al} \gg 1$), then the fixed
points of Eqs.~(\ref{eq:dqdt_toy}--\ref{eq:dfdt_toy}) are slightly modified CSs.
To leading order, all of the CS obliquities $\theta_{\rm cs}$ are unchanged
while the azimuthal angle $\phi_{\rm cs}$ for each CS now satisfies
\begin{equation}
    \sin \phi_{\rm cs} = \frac{\sin\theta_{\rm cs}}{\sin I \abs{g}t_{\rm al}}.
        \label{eq:mcs_shift}
\end{equation}
We can see that if $t_{\rm al}$ is longer than some critical alignment
timescale $t_{\rm al, c}$, given by
\begin{align}
    t_{\rm al, c} &\equiv \frac{1}{\abs{g}\sin I},\label{eq:mcs_shift_crit}
\end{align}
then Eq.~\eqref{eq:mcs_shift} will always have solutions for $\phi_{\rm cs}$,
and the alignment torque does not change the number of fixed points of the
system. If $t_{\rm al}$ is decreased below $t_{\rm al, c}$, CS2 and CS4 cease to
be fixed points when $\eta \ll 1$ \citep[as first noted in][]{fabrycky_otides},
as $\theta_{\rm cs} \approx 90^\circ$ for these (see Fig.~\ref{fig:cs_locs}). On
the other hand, the other CSs have small $\sin \theta_{\rm cs}$ and are only
slightly perturbed. Figure~\ref{fig:mcs} shows the obliquity and azimuthal
angles for each of the CSs in the $\eta \ll 1$ case obtained via numerical
root finding of Eqs.~(\ref{eq:dqdt_toy}--\ref{eq:dfdt_toy}), where it can be
seen that CS2 and CS4 collide and annihilate when $t_{\rm al}$ reaches $t_{\rm
al, c}$. The phase shift $\phi_{\rm cs}$ for CS2 and CS4 for $t_{\rm al} >
t_{\rm al, c}$ can be predicted to good accuracy using Eq.~\eqref{eq:mcs_shift}
using $\theta_{\rm cs} \approx \pi/2 - \eta \cos (I) \approx 79^\circ$
\citep{su2020} for both, shown as the dashed lines in the bottom panel of
Fig.~\ref{fig:mcs}. For the remainder of this section, we will consider the case
where $t_{\rm al} \gg t_{\rm al, c}$ and the CSs only differ slightly from their
unperturbed locations.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/0_stab.png}
    \caption{Modified CS obliquities (top) and azimuthal angles (bottom) for $I
    = 20^\circ$ and $\eta = 0.2$, where the CS1 and CS3 obliquities have been
    offset (labeled in top legend) to improve clarity of the plot. In both
    panels, the solid lines give the result when applying a numerical root
    finding algorithm to the full equations of motion,
    Eqs.~(\ref{eq:dqdt_toy}--\ref{eq:dfdt_toy}), while the dotted lines in the
    bottom panel give the CS2 and CS4 azimuthal angles according to
    Eq.~\eqref{eq:mcs_shift}. At $\abs{gt_{\rm al} \sin I} = 1$, CS2 and CS4
    collide and annihilate (Eq.~\ref{eq:mcs_shift_crit}).}\label{fig:mcs}
\end{figure}

We next seek to characterize the stability of small perturbations about each of
the CSs in the presence of the weak alignment torque. We can linearize
Eqs.~(\ref{eq:dqdt_toy}--\ref{eq:dfdt_toy}) about a shifted CS, yelding
\begin{align}
    \rd{}{t}\begin{bmatrix}
        \theta\\ \phi
    \end{bmatrix} &= \begin{bmatrix}
        -\frac{\cos \theta}{t_{\rm al}} &
        -g\sin I \cos \phi \\
        \alpha \sin \theta + g\frac{\sin I \cos \phi}{\sin^2\theta} &
        0
    \end{bmatrix}_{\rm cs}\begin{bmatrix}
        \Delta \theta \\ \Delta \phi
    \end{bmatrix},\label{eq:dsdt_hessian}
\end{align}
where the cs subscript indicates evaluating at a CS, $\Delta \theta = \theta -
\theta_{\rm cs}$, and $\Delta \phi = \phi - \phi_{\rm cs}$. The eigenvalues
$\lambda$ of Eq.~\eqref{eq:dsdt_hessian} satisfy the equation
\begin{equation}
    0 = \p{\lambda + \frac{\cos \theta_{\rm cs}}{t_{\rm al}}}\lambda
        - \lambda_0^2,\label{eq:lambda_orig}
\end{equation}
where
\begin{equation}
    \lambda_0^2 \equiv \p{\alpha
        \sin \theta_{\rm cs} + g\sin I \csc^2\theta_{\rm cs}\cos \phi_{\rm cs}}
            \p{- g \sin I \cos \phi_{\rm cs}}\label{eq:def_l0_sq}.
\end{equation}
When $t_{\rm al}$ is large, we can simplify Eq.~\eqref{eq:lambda_orig} to
\begin{equation}
    \lambda \approx -\frac{\cos \theta_{\rm cs}}{t_{\rm al}}
        \pm \sqrt{\lambda_0^2}. \label{eq:lambda_approx}
\end{equation}
The stability depends only on the real part of $\lambda$ in
Eq.~\eqref{eq:lambda_approx}. $\lambda_0^2$ is a generalization of Eq.~(A4)
in Paper I and generally has the same behavior: it is negative for CSs
1--3 and positive for CS4, as shown in Fig.~\ref{fig:lambda_full}. Thus, CS4 is
always unstable, as there will always be at least one positive solution for
$\lambda$, and the stability of CSs 1--3 are solely determined by the sign of
$\cos \theta_{\rm cs}$. Using Fig.~\ref{fig:cs_locs}, we conclude that CS1 and
CS2 are stable and attracting while trajectories near CS3 are driven away by the
alignment torque. These calculations justify results long used in CS literature
\citep[e.g.][]{ward1975tidal, fabrycky_otides}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/2_lambdas_full.png}
    \caption{$\lambda_0^2$ (Eq.~\ref{eq:def_l0_sq}) as a function of $\eta$ for
    the four CSs, for three different values of the shift in $\phi_{\rm cs}$
    (e.g.\ for $\Delta \phi_{\rm cs} = 60^\circ$, $\phi_{\rm cs} = 120^\circ$
    for CS2 and $\phi_{\rm cs} = 60^\circ$ for CSs 1, 3, and 4). The values of
    $\Delta \phi_{\rm cs}$ are labeled. }\label{fig:lambda_full}
\end{figure}

\subsection{Spin Obliquity Evolution Driven by Alignment
Torque}\label{ss:toy_outcomes}

With the above results, we are equipped to ask questions about the dynamics of
Eqs.~(\ref{eq:dqdt_toy}--\ref{eq:dfdt_toy}): what is the long term behavior for
a general initial $\uv{s}$? If $\eta > \eta_{\rm c}$, then the only stable
final spin state is CS2, and all initial conditions will evolve asympotically
towards it. As such, we consider only $\eta < \eta_{\rm c}$, where an initial
condition can asymptotically evolve towards either CS1 or CS2. Additionally, we
restrict our attention to the regime where the alignment torque is weak and CS2
remains stable; specifically, we use $\abs{g}t_{\rm al}= 10^{3}$. We
numerically integrate Eqs.~(\ref{eq:dqdt_toy}--\ref{eq:dfdt_toy}) for many
random initial conditions uniformly distributed in $\p{\cos \theta, \phi}$ and
record the nearest CS for each integration after $10t_{\rm al}$. In
Fig.~\ref{fig:toy_phop}, we show the results of this procedure for $\eta = 0.2$,
and $I = 20^\circ$. It is clear that initial conditions in zone I evolve into
CS1, those in zone II evolve into CS2, while those in zone III have a
probabilistic outcome. We aim to understand each of these in turn:
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/0_eta/3stats3_5_0_2.png}
    \caption{Plot illustrating the asymptotic behavior of initial conditions for
    $\eta = 0.2$ and $I = 20^\circ$. Each dot represents an initial spin
    orientation, and the coloring of the dot indicates which Cassini State
    (legend) the system asymptotes towards.}\label{fig:toy_phop}
\end{figure}

For initial conditions in zone I, the spin circulates, and $\dot{\theta}$ is
negative everywhere during the cycle. Thus, for initial conditions in zone I,
$\theta$ decreases until the trajectory has converged to CS1. This is
intuitively reasonable, as CS1 is stable.

For initial conditions in zone II, our stability calculation in
Section~\ref{ss:tidal_equils} shows that initial conditions sufficiently near CS2
will converge to CS2. In fact, this result can be extended to all initial
conditions inside the separatrix, shown in Appendix~\ref{app:cs_stab2}.

For initial conditions in zone III, since there are no stable CSs in zone III,
the system must evolve through the separatrix to reach one of either CS1 or CS2.
The outcome of the separatrix encounter is effectively probabilistic and
determines the final CS\@. Intuitively, this can be understood as probabilistic
resonance capture, as first studied in the seminal work of \citet{henrard1982}:
since $\eta \ll \eta_{\rm c}$, $\alpha \gg -g$, but $\alpha \cos \theta$ can
become commensurate with $-g$ if $\cos \theta$ becomes small. This is achieved
as $\theta$ evolves from an initially retrograde obliquity through $90^\circ$
towards $0^\circ$ under the influence of the dissipative term in
Eq.~\eqref{eq:dqdt_toy}.

While similar in behavior to previous studies of probabilistic resonance capture
\citep{henrard1982, su2020}, the underlying mechanism is different: in these
previous studies, the phase space structure itself evolves and causes systems to
transition among phase space zones, while in the present scenario, a
non-Hamiltonian, dissipative perturbation causes systems to transition among
unchanging phase space zones. In the following section, we present a calculation
to understand analytically the probability distribution of outcomes upon
separatrix encounter in this new scenario.

\subsubsection{Analytical Calculation of Resonance Capture Probability}

Before discussing quantitative calculations, we first present a graphical
understanding of the separatrix encounter process.
Figure~\ref{fig:toy_hop_manifolds} shows how the perturbative alignment torque
generates the two outcomes upon separatrix encounter, a zone III-II or a zone
III-I transition. The critical trajectories in Fig.~\ref{fig:toy_hop_manifolds}
are calculated numerically by integrating a point infinitesimally close to CS4
forward and backward in time. In the absence of the alignment torque, these
trajectories would evolve along the separatrix, but in the presence of the
alignment torque, they are perturbed slightly and cease to overlap. It can be
seen in Fig.~\ref{fig:toy_hop_manifolds} that this splitting opens a path from
zone III into both zones I and II\@.

To understand this process more concretely, as well as associate probabilities
to the two possible outcomes, it is important to be quantitative. The correct
approach is to consider the evolution of the value of the \emph{unperturbed}
Hamiltonian (Eq.~\ref{eq:H}) as the spin evolves due to the alignment torque. A
point in zone III evolves such that $H$ is increasing until $H \approx H_{\rm
sep}$ where $H_{\rm sep}$ is the value of $H$ along the separatrix, given by
\begin{align}
    H_{\rm sep} &\equiv H\p{\cos \theta_{\rm 4}, \phi_{\rm 4}},\nonumber\\
        &\approx g\sin I + \frac{g^2}{2\alpha}\cos^2 I +
            \mathcal{O}\p{\eta^2},\label{eq:def_Hsep}
\end{align}
where $\theta_4$ and $\phi_4$ are the coordinates of CS4. As the system evolves
closer to the separatrix, the change in $H$ over each circulation cycle can be
approximated by $\Delta H_-$, the change in $H$ along $\mathcal{C}_-$ (see
Fig.~\ref{fig:1contours}). In general, we define the quantities $\Delta H_{\pm}$
\begin{equation}
    \Delta H_{\pm} \equiv \oint\limits_{\mathcal{C}_{\pm}}
        \rd{H}{t}\;\mathrm{d}t.\label{eq:def_dHpm}
\end{equation}
This can be simplified by using:
\begin{align}
    \rd{H}{t} &=
            \pd{H}{(\cos \theta)}\rd{(\cos \theta)}{t}
            + \pd{H}{\phi}\rd{\phi}{t},\nonumber\\
        &= \p{\rd{(\cos\theta)}{t}}_{\rm tide} \rd{\phi}{t},\\
    \Delta H_{\pm} &= \mp\frac{1}{t_{\rm al}}
        \int\limits_0^{2\pi} \sin^2\theta\;\mathrm{d}\phi.
\end{align}
In the first line, Hamilton's equations ensure that the non-tidal contributions
to $\rdil{H}{t}$ vanish. Thus, if we evaluate $H$ every time that a trajectory
originating in zone III crosses $\phi = 0$, we see that it will initially be $<
H_{\rm sep}$ and increase for each circulation cycle until the system encounters
the separatrix. At the beginning of the separatrix-crossing orbit, the initial
value of $H$, denoted $H_{\rm i}$, must be greater than $H_{\rm sep} - \Delta H$
to encounter the separatrix on the current orbit, and must be less than $H_{\rm
sep}$, otherwise the trajectory will be in zone II already. We thus have
\begin{equation}
    H_{\rm i} \in \s{ H_{\rm sep} - \Delta H_-,  H_{\rm sep} }
        \label{eq:Hi_range},
\end{equation}
The values of $\cos \theta$ corresponding to the lower and upper bounds on this
range are shown as the black and red dots on the left of Fig.~\ref{fig:6equils}
respectively.

During the separatrix-crossing orbit, the trajectory first evolves
approximately along $\mathcal{C}_-$ and then along $\mathcal{C}_+$, after which
the final value of $H$, denoted $H_{\rm f}$ is approximately equal to
\begin{equation}
    H_{\rm f} = H_{\rm i} + \Delta H_+ + \Delta H_-.
\end{equation}
There can then be two outcomes depending on the value of $H_{\rm f}$:
\begin{itemize}
    \item If $H_{\rm f} > H_{\rm sep}$, then the trajectory enters the
        separatrix, following the red shaded region in Fig.~\ref{fig:6equils},
        and executes a zone III $\to$ II transition.

    \item If $H_{\rm f} < H_{\rm sep}$, then the trajectory exits the separatrix
        above CS4, following the yellow shaded region in Fig.~\ref{fig:6equils}
        and executes a zone III $\to$ I transition.
\end{itemize}
These two possibilites can be re-expressed in terms of $H_{\rm i}$: we find that
if $H_{\rm i}$ is in the interval $\s{H_{\rm sep} - \Delta H_-, H_{\rm sep} -
\Delta H_{-} - \Delta H_+}$, then the system executes a III $\to$ I transition,
and if it is in the interval $\s{H_{\rm sep} - \Delta H_- - \Delta H_+, H_{\rm
sep}}$, then the system executes a III $\to$ II transition. We see that there is
a critical value of $H_{\rm i}$ in the interval given by
Eq.~\eqref{eq:Hi_range}, where $H_{\rm i} = H_{\rm sep} - \Delta H_{-} - \Delta
H_+$, that separates the two possible outcomes of the separatrix encounter
within the interval given by Eq.~\eqref{eq:Hi_range}. The value of $\cos \theta$
for which $H$ is equal to $H_{\rm sep} - \Delta H_{-} - \Delta H_+$ is shown as
the blue dot on the left of Fig.~\ref{fig:6equils}. Finally, if the alignment
torque is weak, then $\Delta H_- \sim \mathcal{O}\p{t_{\rm al}^{-1}}$ is small
compared to any variation in the value of $H$ (e.g.\ when changing $\phi_0$ or
$\theta_0$ by a small amount), and $H_{\rm i}$ can be effectively considered as
randomly chosen from a uniform distribution over $\s{H_{\rm sep} - \Delta H_-,
H_{\rm sep}}$. As a consequence we obtain the probability of a III $\to$ II
transition:
\begin{equation}
    P_{\rm III \to II} = \frac{\Delta H_- + \Delta H_+}{\Delta H_-}.
        \label{eq:def_P32_toy}
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/0_eta/6manifolds0_20.png}
    \caption{Plot illustrating the probabilistic origin of separatrix capture
    for $\eta = 0.2$, $I = 20^\circ$, and $\abs{gt_{\rm al}} = 10^{3}$. Orange
    regions converge to CS1, and green to CS2, while CS4 is labeled with the
    purple dots. The boundaries separating the CS1 and CS2-approaching regions
    consist of the critical trajectories (labeled in the legend) that, when
    evolved either forwards or backwards in time (superscripts in the legend
    labels), asymptotically approach CS4 where either $\phi = 0^\circ$ or $\phi
    = 360^\circ$ (subscripts in the legend labels). Within zone III, the regions
    of phase space reaching CS1 and CS2 both become very thin (shown
    qualitatively as the light green lines of decreasing width), reflecting the
    fact that the outcome for a particular initial condition can be approximated
    as probabilistic sufficiently far from the separatrix. Labeled in the blue
    and black dots at the left edge of the plot are the two backwards-in-time
    critical trajectories intersect $\phi = 0$, and are where $H$ is equal to
    $H_{\rm sep} - \Delta H_- - \Delta H_+$ and $H_{\rm sep} - \Delta H_-$
    respectively (Eqs.~\ref{eq:def_Hsep} and~\ref{eq:def_dHpm}).
    }\label{fig:toy_hop_manifolds}
\end{figure}

To actually evaluate Eq.~\eqref{eq:def_P32_toy}, we can use the parameterization
for the separatrix given in Paper I for $\eta \ll 1$:
\begin{equation}
    \p{\cos \theta}_{\mathcal{C}_{\pm}} \approx
        \eta \cos I \pm \sqrt{2\eta\sin I\p{1 - \cos \phi}}.
        \label{eq:sep_theta}
\end{equation}
It can then be shown that
\begin{align}
    \Delta H_- &\approx \frac{2\pi}{t_{\rm al}}\p{1
        - 2\eta \sin I} + \mathcal{O}(\eta^{3/2}),\\
    \Delta H_+ + \Delta H_- &\approx
        \frac{32 \eta^{3/2}\cos I \sqrt{\sin I}}{t_{\rm al}}
            + \mathcal{O}\p{\eta^{5/2}},\\
    P_{\rm III \to II} &\approx
        \frac{16 \eta^{3/2} \cos I \sqrt{\sin I}}{\pi
            \p{1  - 2\eta \sin I}}.\label{eq:P32_toy}
\end{align}

To directly compare Eq.~\eqref{eq:P32_toy} with numerical simulation, we perform
numerical integrations of Eqs.~(\ref{eq:dqdt_toy}--\ref{eq:dfdt_toy}) while
restricting our attention only initial conditions in zone III\@.
In Fig.~\ref{fig:1hist_toy}, we display Eq.~\eqref{eq:P32_toy} alongside the
computed $P_{\rm III \to II}$ using $1000$ initial conditions in zone III for
each of $60$ values of $\eta$. Excellent agreement is observed.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/3_toy3/1hist_toy.png}
    \caption{Plot of $P_{\rm III \to II}$ as a function of $\eta$ for the
    constant alignment torque model considered in Section~\ref{s:toy_model}. For
    each $\eta$, $1000$ initial $\theta_0$ in zone III are evolved until just
    after separatrix encounter, where the outcome of the encounter is recorded.
    Shown in red is Eq.~\eqref{eq:P32_toy}.}\label{fig:1hist_toy}
\end{figure}

Finally, we remark that the calculation above is just an of \emph{Melnikov's
Method} \citep{g_and_h}. Melnikov's Method is a general calculation that gives
the degree of splitting of a ``homoclinic orbit'' (here, the separatrix) induced
by a small, possibly time-dependent, perturbation. The explicit connection
between the evolution of the unperturbed Hamiltonian and distances between
curves in phase space (as seen in Fig.~\ref{fig:6equils}) is provided by
Melnikov's Method.

\section{Spin Evolution with Weak Tidal Friction}\label{s:full_tide_prob}

\subsection{Tidal Model: Equilibrium Tides}\label{ss:weaktide}

Having understood the effect of introducing an additional dissipative torque to
the Colombo's Top system, we now implement a realistic model of tidal
dissipation. We use the weak friction theory of equilibrium tides
\citep{lai2012}. In this model, tides cause both the spin orientation $\uv{s}$
and frequency $\Omega_{\rm s}$ to evolve on the characteristic tidal timescale
$t_{\rm s}$ following:
\begin{align}
    \p{\rd{\uv{s}}{t}}_{\rm tide} &= \frac{1}{t_{\rm s}}
                \s{\frac{2n}{\Omega_{\rm s}} - \p{\uv{s} \cdot \uv{l}}}
                    \uv{s} \times \p{\uv{l} \times \uv{s}}\label{eq:dsdt_tide},\\
    \frac{1}{\Omega_{\rm s}}\p{\rd{\Omega_{\rm s}}{t}}_{\rm tide}
        &= \frac{1}{t_{\rm s}} \s{\frac{2n}{\Omega_{\rm s}}\p{\uv{s} \cdot
            \uv{l}} - 1 - \p{\uv{s} \cdot \uv{l}}^2},\label{eq:dWsdt_tide}
\end{align}
where $t_{\rm s}$ is given by
\begin{equation}
    \frac{1}{t_{\rm s}} \equiv \frac{L}{2S} \frac{\Omega_{\rm
        s}}{2n}\frac{3k_2}{Q}\p{\frac{M_\star}{m}}\p{\frac{R}{a}}^5 n.
\end{equation}
Here, $L = ma^2n$ and $S = kmR^2 \Omega_{\rm s}$ are the orbital and spin
angular momenta of the inner planet, respectively. We neglect orbital evolution
in this section since the time scale is longer than $t_{\rm s}$ by a factor of
$\sim L / S \gg 1$, and so $t_{\rm s}$ is a constant. We will continue mostly
consider the case where tidal dissipation is slow, i.e.\ $\abs{g}t_{\rm s} \gg
1$. The full equations of motion including weak tidal friction can be written in
component form as
\begin{align}
    \rd{\theta}{t} &= g\sin I \sin \phi -
        \frac{1}{t_{\rm s}}\sin \theta\p{\frac{2n}{\Omega_{\rm s}} - \cos \theta}
            ,\label{eq:ds_fullq}\\
    \rd{\phi}{t} &= -\alpha\cos\theta
        - g\p{\cos I + \sin I \cot \theta \cos \phi}\label{eq:ds_fullphi},\\
    \frac{1}{\Omega_{\rm s}}\rd{\Omega_{\rm s}}{t}
        &= \frac{1}{t_{\rm s}} \s{\frac{2n}{\Omega_{\rm s}} \cos \theta
            - \p{1 + \cos^2\theta}}\label{eq:ds_fulls}.
\end{align}
The conditions for which the two tidal terms vanish are given respectively by:
\begin{align}
    \dot{\theta}_{\rm tide} = 0 &: \quad \frac{2n}{\Omega_{\rm s}} = \cos \theta,
        \label{eq:weaktide_dqzero}\\
    \dot{\Omega}_{\rm s} = 0 &: \quad \frac{n}{\Omega_{\rm s}}
        = \frac{1 + \cos^2\theta}{2\cos \theta}.\label{eq:weaktide_dWszero}
\end{align}

To understand the long-term behaviors of this system, we first consider its
behavior near a CS\@. Specifically, we wish to understand whether initial
conditions near a CS stay near the CS as the evolution of $\Omega_{\rm s}$
causes the CSs to evolve. We first note that the evolution of
$\Omega_{\rm s}$ does not drive spins towards or away from CSs: As long as it
evolves sufficiently slowly (adiabatically, Paper I), conservation of
phase space area ensures that trajectories will remain at a roughly fixed
distance to stable equilibria of the system. Thus, Eq.~\eqref{eq:dsdt_tide}
alone determines whether a point evolves towards or away from a nearby CS as
$\Omega_{\rm s}$ evolves. Then, evaluating Eq.~\eqref{eq:lambda_approx} with
$t_{\rm al} = t_{\rm s} / \p{2n / \Omega_{\rm s} - \cos \theta}$ (compare
Eqs.~\ref{eq:dsdt_tide_toy} and~Eq.~\ref{eq:dsdt_tide}), we see that CS2 is
still always stable, while CS1 is becomes unstable for $\Omega_{\rm s} > 2n$.

Using this result, we can then identify the long-term equilibria of the system
when $\Omega_{\rm s}$ is also evolving: these equilibria of the system must both
satisfy $\dot{\Omega}_{\rm s} = 0$ and be a CS that is stable in the presence of
tidal dissipation (i.e.\ satisfying $\rdil{\uv{s}}{t} = 0$).
The panels of Fig.~\ref{fig:6equils} describes the evolution of
Eqs.~(\ref{eq:dsdt_tide}--\ref{eq:dWsdt_tide}) qualitatively in the coordinates
$(\Omega_{\rm s}, \theta)$, along with the locations of CS1 and CS2, for a few
different system parameters. The circled points in the panels of
Fig.~\ref{fig:6equils} satisfy the criteria to be long-term equilibria, and we
call them \emph{tidal Cassini Equilibria} (tCE). We number tCE1 and tCE2 the tCE
that are in CSs 1 and 2 respectively. The obliquities of the tCE depend on the
system architecture, which can be quantified using the parameter
\begin{align}
    \eta_{\rm sync} &\equiv \big(\eta\big)_{\Omega_{\rm s} = n}
        = \eta \frac{\Omega_{\rm s}}{n}\nonumber\\
        &= \frac{1}{2}\frac{k}{k_{\rm q}}
            \frac{m_{\rm p}m}{M_\star^2}
            \p{\frac{a}{a_{\rm p}}}^3 \p{\frac{a}{R}}^3\cos(I).
            \label{eq:def_etasync}
\end{align}
The tCE obliquity as a function of $\eta_{\rm sync}$ are shown for $I =
20^\circ$ in the top panel of Fig.~\ref{fig:probs20} and for $I = 5^\circ$ in
the top panel of Fig.~\ref{fig:probs5}. In fact, an analytical expression for
the tCE2 obliquity for $\eta_{\rm sync} \ll \eta_{\rm c}$ can be computed using
Eqs.~(\ref{eq:weaktide_dWszero}--\ref{eq:def_etasync}) and that $\cos \theta_2
\approx \eta \cos I$ to yield:
\begin{align}
    \cos \theta_{\rm tCE2} &\approx \sqrt{\frac{\eta_{\rm sync} \cos I}{2}},
        &
    \frac{\Omega_{\rm s, tCE2}}{n} &\approx
        \sqrt{2\cos I \eta_{\rm sync}}. \label{eq:def_tce2_approx}
\end{align}
This approximation for $\theta_{\rm tCE2}$ is shown as the black dashed line in
the top panels of Fig.~\ref{fig:probs20} and~\ref{fig:probs5}, where good
agreement is observed.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/6equils.png}
    \caption{Schematic depiction of the effect of tidal friction on the planet's
    spin for $I = 20^\circ$ and three values of $\eta_{\rm sync}$ (labeled). The
    black and blue lines denote where the tidal $\dot{\Omega}_{\rm s}$ and
    $\dot{\theta}$ change signs
    (Eqs.~\ref{eq:weaktide_dqzero}--\ref{eq:weaktide_dWszero}). The orange and
    green lines give the CS1 and CS2 obliquities respectively, which are the two
    CSs that are stable under the effect of tidal dissipation. Note that when
    $\dot{\theta}_{\rm tide} > 0$, CS1 becomes unstable, denoted by the dashed
    orange line. The points that are both CSs and satisfy $\dot{\Omega}_{\rm s}
    = 0$ are the tidal Cassini Equilibria (tCE), which are circled and labeled.
    }\label{fig:6equils}
\end{figure}

There are two important conditions that can change the existence and stability
of the tCE\@. First, if $\eta_{\rm sync} > \eta_{\rm c}$, then tCE1 will not
exist, as can be seen in the bottom panel of
Fig.~\ref{fig:6equils}\footnote{Strictly speaking, $\eta_{\rm sync}$ can be
slightly smaller than $\eta_{\rm c}$, as the planet's spin is slightly
subsynchronous at tCE1 when $\eta \approx \eta_{\rm c}$, see the middle panel of
Fig.~\ref{fig:6equils}.}. Secondly, tCE2 may not be stable if the tidal phase
shift is too large. Applying the results of Section~\ref{ss:tidal_equils}, we
find that tCE2 is stable as long as $t_{\rm s} \geq t_{\rm s, c}$ where
\begin{equation}
    t_{\rm s, c} \equiv \frac{\sin \theta_{\rm tCE2}}{\abs{g} \sin
            I}\p{\frac{2n}{\Omega_{\rm s, tCE2}} - \cos \theta_{\rm tCE2}}.
\end{equation}
The tCE2 subscripts denote evaluation at tCE2. When $\eta_{\rm sync} \ll
\eta_{\rm c}$, we can use Eq.~\eqref{eq:def_tce2_approx} to further simplify:
\begin{equation}
    t_{\rm s, c}\approx \frac{1}{\abs{g}\sin I}\sqrt{\frac{2}{\eta_{\rm sync}
        \cos I}}. \label{eq:def_ts_crit}
\end{equation}

\subsection{Spin Obliquity Evolution as a Function of Initial Spin Orientation}

With this result, we can now consider the final fate of the inner planet's spin.
We assume that the planet is initially rotating supersynchronously and adopt the
fiducial initial spin frequency $\Omega_{\rm s, i} = 10n$. The final results are
not sensitive to the specific initial spin as long as $\Omega_{\rm s, i} \gg n$.
Then, for a given initial $\theta_0$ and $\phi_0$, the final outcome (either
tCE1 and tCE2) of the system can be calculated by direct integration of
Eqs.~(\ref{eq:dsdt_rot},~\ref{eq:dsdt_tide}--\ref{eq:dWsdt_tide}). In
Fig.~\ref{fig:Hhists_0_06}, we show the final outcome for many randomly chosen
$\theta_0$ and $\phi_0$ for $\eta_{\rm sync} = 0.06$ and $I = 20^\circ$. We see
that tCE1 is generally reached for spins initially in zone I, tCE2 is generally
reached for spins initially in zone II, and a probabilistic outcome is observed
for spins initially in zone III, similarly to Fig.~\ref{fig:toy_phop}.
Figures~\ref{fig:Hhists_0_20} and~\ref{fig:Hhists_0_70} show the same results
but for $\eta_{\rm sync} = 0.2$ and $\eta_{\rm sync} = 0.5$. As $\eta_{\rm
sync}$ is increased, more initial conditions reach tCE2. This is both because
there are more systems initially in zone II and because systems initially
in zone III have a higher probability of executing a III $\to$ II transition
upon separatrix encounter. Note also that in Fig.~\ref{fig:Hhists_0_70}, even
initial conditions in zone I are able to reach tCE2; we comment on the origin of
this behavior in the next section.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_06_20.png}
    \caption{\emph{Left:} Each dot indicates which tCE a given initial condition
    $(\theta_{\rm i}, \phi_{\rm i})$ evolve towards (labeled in legend), for
    $\eta_{\rm sync} = 0.06$ and $I = 20^\circ$. The planet's initial spin is
    chosen to be $\Omega_{\rm s, i} = 10n$, and the three zones of phase space
    (see Fig.~\ref{fig:1contours}) for the initial conditions are labeled,
    separated by the separatrix (black line). Note that initial conditions in
    zone I evolve towards tCE1, initial conditions in zone II evolve towards
    tCE2, and initial conditions in zone III have an effectively probabilistic
    outcome. \emph{Right:} Histogram of which tCE a given initial obliquity
    $\theta_{\rm i}$ evolves towards, averaged over $\phi_{\rm
    i}$.}\label{fig:Hhists_0_06}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_20_20.png}
    \caption{Same as Fig.~\ref{fig:Hhists_0_06} but for $\eta_{\rm sync} =
    0.2$.}\label{fig:Hhists_0_20}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5Hhists0_50_20.png}
    \caption{Same as Fig.~\ref{fig:Hhists_0_06} but for $\eta_{\rm sync} =
    0.5$. Note that even points above the separatrix can evolve towards tCE2
    here.}\label{fig:Hhists_0_70}
\end{figure}

\subsubsection{Analytical Calculation of Resonance Capture Probability
}\label{ss:phop_weaktide}

Even when including the evolution of $\Omega_{\rm s}$, and therefore the
spin-orbit precession frequency $\alpha$, the probabilities of the III $\to$ I
and III $\to$ II transitions upon separatrix encounter can still be computed.
The calculation resembles that presented in Section~\ref{ss:toy_outcomes} but is
more involved. We describe the calculation below.

In Section~\ref{ss:toy_outcomes}, we found that the evolution of $H$, the value
of the unperturbed Hamiltonian, allowed us to calculate the probabilities of the
various outcomes of separatrix encounter. Specifically, the outcome upon
separatrix encounter is determined by the value of $H$ at the start of the
separatrix-crossing orbit relative to $H_{\rm sep}$, the value of $H$ along the
separatrix. However, when the spin is also evolving, $H_{\rm sep}$ is also
changing during the separatrix-crossing orbit, and the discussion in
Section~\ref{ss:toy_outcomes} must be generalized to account for this. Instead
of focusing on the evolution of $H$ along a trajectory, we instead follow the
evolution of
\begin{equation}
    K \equiv H - H_{\rm sep}.
\end{equation}
Note that $K > 0$ inside the separatrix, and $K < 0$ outside. Then, the outcome
of the separatrix-crossing orbit is largely the same as discussed in
Section~\ref{ss:toy_outcomes}. First, we must compute the change in $K$ along
the legs of the separatrix. We define $\Delta K_{\pm}$ by generalizing
Eq.~\eqref{eq:def_dHpm} in the natural way:
\begin{align}
    \Delta K_{\pm} &= \oint_{\mathcal{C}_{\pm}} \rd{H}{t}
        - \rd{H_{\rm sep}}{t}\;\mathrm{d}t.\label{eq:def_dK_weaktide}
\end{align}
Here, however, note that the contours $\mathcal{C}_{\pm}$ depends on the value
of $\Omega_{\rm s}$ at separatrix encounter. Since there is no closed form
solution for $\Omega_{\rm s}(t)$, the probabilities of the various outcomes
cannot be expressed as a simple function of the initial conditions.

Continuing the argument presented in Section~\ref{ss:toy_outcomes}, we consider
the outcome of the separatrix-crossing orbit as a function of $K_i$, the value
of $K$ at the start ($\phi = 0$) of the separatrix-crossing orbit. We find that
if $K_i > -\Delta K_+ - \Delta K_-$, then the system undergoes a III $\to$ II
transition and eventual evolution towards tCE2; while if $-\Delta K_- < K_i <
-\Delta K_- - \Delta K_+$, then the system undergoes a III $\to$ I transition
and ultimate evolution towards tCE1. Thus, we find that the probability of a III
$\to$ II transition is given by
\begin{equation}
    P_{\rm III \to II} = \frac{\Delta K_+ + \Delta K_-}{\Delta
        K_-}.\label{eq:def_pc_weaktide}
\end{equation}
Again, since $\Delta K_{\pm}$ are evaluated at resonance encounter, and
$\Omega_{\rm s}$ is evolving, there is no way to express $\Delta K_{\pm}$ as a
closed form of initial conditions. In fact, since many resonance encounters
occur when $\eta$ is substantial, even an approximate calculation of $\Delta
K_{\pm}$ using Eq.~\eqref{eq:sep_theta} is inaccurate, and we instead calculate
$\Delta K_{\pm}$ along the numerically-calculated $\Delta K_{\pm}$ for arbitrary
$\eta$. Note that Eqs.~(\ref{eq:def_dK_weaktide},~\ref{eq:def_pc_weaktide}) are
equivalent to the separatrix capture result of \citet{henrard1982} when
$\dot{\theta}_{\rm tide} = 0$ \citep{henrard1987}. Here, we have argued that
this classic calculation can be unified with the calculation given in
Section~\ref{ss:toy_outcomes} to give an accurate prediction of separatrix
encounter outcome probabilities in the presence of both a dissipative
perturbation and a parametric evolution of the Hamiltonian.

To validate the accuracy of Eq.~\eqref{eq:def_pc_weaktide}, we can compare with
direct numerical integration of
Eqs.~(\ref{eq:dsdt_rot},~\ref{eq:dsdt_tide}--\ref{eq:dWsdt_tide}) for many
initial conditions while evaluating $P_{\rm III \to II}$ (and thus also
obtaining $P_{\rm III \to I} = 1 - P_{\rm III \to II}$) for each simulation at
the moment it encounters the separatrix, if it does so. If the theory is
correct, the total numbers of systems converging to each of tCE1 and tCE2 should
be equal to those predicted by the sums of the calculated probabilities. In
Fig.~\ref{fig:pc_fits_0_06}, we show the agreement of this semi-analytic
procedure with the numerical results displayed earlier in the right panel of
Fig.~\ref{fig:Hhists_0_06}. Good agreement is observed.
Figs.~\ref{fig:pc_fits_0_20} shows the same for Figs.~\ref{fig:Hhists_0_20}.
Mostly satisfactory agreement is observed. Thus, we conclude that the outcomes
of separatrix encounter are accurately predicted by
Eq.~\eqref{eq:def_pc_weaktide}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5pc_fits0_06_20.png}
    \caption{Comparison of the fraction of systems converging to tCE2 obtained
    via numerical simulation (red dots) and obtained via a semi-analytic
    calculation (blue line) for $\eta_{\rm sync} = 0.06$ and $I = 20^\circ$ (see
    right panel of Fig.~\ref{fig:Hhists_0_06}). The semi-analytic calculation is
    performed by numerically integrating
    Eqs.~(\ref{eq:dsdt_rot},~\ref{eq:dsdt_tide}--\ref{eq:dWsdt_tide}) on a grid
    of initial conditions uniform in $\cos \theta_{\rm i}$ and $\phi_{\rm i}$
    until the system reaches the separatrix, then calculating the probability of
    reaching tCE2 for each integration using
    Eq.~\eqref{eq:def_pc_weaktide}.}\label{fig:pc_fits_0_06}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5pc_fits0_20_20.png}
    \caption{Same as Fig.~\ref{fig:pc_fits_0_06} but for $\eta_{\rm sync} =
    0.2$, corresponding to the right panel of
    Fig.~\ref{fig:Hhists_0_20}.}\label{fig:pc_fits_0_20}
\end{figure}

With the above calculation, we can understand why even initial conditions in
zone I can converge to tCE2. As long as the initial spin is sufficiently large
($\geq 2n$), Eq.~\eqref{eq:dsdt_tide} shows that sufficiently large initial
obliquities are \emph{increased} by weak tidal friction (see the bottom right
region of Fig.~\ref{fig:6equils}). This increase means that even initial
conditions in zone I can sometimes be driven by tidal dissipation into the
separatrix. When this is the case, it can be shown that both $\Delta K_+ > 0$
and $\Delta K_- > 0$. This implies that both III $\to$ II and I $\to$ II
outcomes are guaranteed, and thus initial conditions in all three zones can
evolve into tCE2.

\subsection{Spin Obliquity Evolution as a Function of Precession Strength
}\label{ss:tce2_etasync}

In the previous section, we considered the outcome as a function of the initial
spin orientation, specified by $\theta_0$ and $\phi_0$. In this section, we
consider the probability of evolution into tCE2 when averaging over a
distribution of initial spin orientations. For simplicity, we just consider
$\uv{s}$ to be isotropically distributed; we discuss the impact of more
physically realistic distributions of $\uv{s}$ in Section~\ref{s:summary}. The
bottom panel of Fig.~\ref{fig:probs20} shows the probability of evolution into
tCE2 for $I = 20^\circ$ as a function of $\eta_{\rm sync}$. It can be seen that
tCE2 is reached with substantial probability ($\sim 50\%$) while it is also
at a substantial obliquity ($\sim 60^\circ$) for $\eta_{\rm sync} \approx 0.4$.
Furthermore, we note that initial conditions other than those beginning in zone
II are responsible for a significant fraction of systems that converge to tCE2.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5probs_20.png}
    \caption{\emph{Top:} Obliquities of the two tCE where $I = 20^\circ$ for a
    range of $\eta_{\rm sync}$ (Eq.~\ref{eq:def_etasync}); the blue dashed line
    denotes the analytical approximation Eq.~\eqref{eq:def_tce2_approx} and is
    only displayed in its regime of validity, $\eta_{\rm sync} \ll \eta_{\rm c}$
    (vertical dashed line; Eq.~\ref{eq:def_etac}). \emph{Bottom:} Total
    probability of ending up in tCE2 (red dots), averaged over an isotropic
    initial spin orientation. The three shaded regions denote the contribution
    of initial conditions in zones I/II/III (labeled) to the total tCE2
    probability in the isotropic average. For example, among systems that
    converge to tCE2 for $\eta_{\rm sync} = 0.06$, more originate in zone III
    than zone II, and none originate in zone I.}\label{fig:probs20}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/1_weaktide/5probs_5.png}
    \caption{Same as Fig.~\ref{fig:probs20} but for $I =
    5^\circ$.}\label{fig:probs5}
\end{figure}

\section{Applications}\label{s:applications}

\subsection{Application to Super Earth + Cold Jupiter Systems
}\label{ss:disc_sehj}

Consider a system consisting of an inner Super-Earth (SE) and an exterior cold
Jupiter (CJ); such systems are expected to be abundant \citep{zhu2018super}. A
phase of giant impacts may occur in the formation of such SEs
\citep{inamdar2015formation, izidoro2017breaking}, which can scramble the
initial obliquity distribution into one resembling the isotropic distribution
considered in Section~\ref{ss:tce2_etasync}. For this system, the spin of the
planet approaches a tCE on the timescale
% 1 / (1.5 * 1e-3 * (Msun / (4 Mearth)) * ((2 Rearth) / (0.4 AU))^3 * 2 * pi / (0.4)^(3/2))
% = 3.34e7
\begin{align}
    \frac{1}{t_{\rm s}} ={}& \frac{1}{3 \times 10^7\;\mathrm{yr}}
            \p{\frac{1}{4k}}
            \p{\frac{2k_2/Q}{10^{-3}}}
            \p{\frac{M_\star}{M_{\odot}}}^{3/2}
            \p{\frac{m}{4M_{\oplus}}}^{-1}\nonumber\\
        &\times \p{\frac{R}{2R_{\oplus}}}^3
            \p{\frac{a}{0.4\;\mathrm{AU}}}^{-9/2}.
\end{align}
This occurs well within the age of most SE-CJ systems. On the other hand,
the orbital evolution of the SE occurs on the timescale \citep{lai2012}
% 1 / (1.5 * 1e-3 * (Msun / (4 Mearth)) * ((2 Rearth) / (0.4 AU))^5 * 2 * pi / (0.4)^(3/2))
% = 7.36e14
\begin{align}
    -\frac{\dot{a}}{a} ={}& \frac{3k_2}{Q}\frac{M_\star}{m}
            \p{\frac{R}{a}}^5n \p{1 - \frac{\Omega_{\rm s}}{n}\cos \theta}\nonumber\\
        ={}& \frac{1}{7 \times 10^{14}\;\mathrm{yr}}
            \p{1 - \frac{\Omega_{\rm s}}{n}\cos \theta}
            \p{\frac{2k_2/Q}{10^{-3}}}
            \p{\frac{M_\star}{M_{\odot}}}^{3/2}\nonumber\\
        &\times \p{\frac{m}{4M_{\oplus}}}^{-1}
            \p{\frac{R}{2R_{\oplus}}}^5
            \p{\frac{a}{0.4\;\mathrm{AU}}}^{-13/2}.
\end{align}
Thus, $a$ does not evolve within the age of the SE-CJ system, and the
approximation where $a$ is held constant is valid. For the same fiducial
parameters, Eq.~\eqref{eq:def_etasync} can be evaluated
% 1/2 * Mjup * (4 * Mearth) / (Msun)^2 * ((0.4 AU) / (2 Rearth))^3 * (0.4 / 5)^3
% = 0.303884
\begin{align}
    \eta_{\rm sync} ={}& 0.303 \cos I
            \p{\frac{k}{k_{\rm q}}}
            \p{\frac{m_{\rm p}}{M_{\rm J}}}
            \p{\frac{m}{4M_{\oplus}}}
            \p{\frac{M_\star}{M_{\odot}}}^{-2}
            \p{\frac{a}{0.4\;\mathrm{AU}}}^{6}\nonumber\\
        &\times \p{\frac{a_{\rm p}}{5\;\mathrm{AU}}}^{-3}
            \p{\frac{R}{2R_{\oplus}}}^{-3}.\label{eq:sehj_etasync}
\end{align}
Here, $M_{\rm J}$ is the mass of Jupiter. We see from Figs.~\ref{fig:probs20}
and~\ref{fig:probs5} that this value of $\eta_{\rm sync}$ gives a high-obliquity
tCE2 with significant probability. Eq.~\eqref{eq:def_ts_crit} shows that tCE2 is
only unstable if $t_{\rm s} \lesssim t_{\rm s, c}$ where
% 1 / (3/8 * ((Mjup^3 * 4 * Mearth) / Msun^4 * (0.4 AU)^12 / ((5 AU)^9 * (2 Rearth)^3))^(1/2) * 2 * pi / (0.4)^(3/2))
% = 2.8189e5
\begin{align}
    \frac{1}{t_{\rm s, c}} ={}& \frac{\sin I \cos^2 I}{3 \times
        10^5\;\mathrm{yr}}
            \frac{k}{k_{\rm q}}
            \p{\frac{m_{\rm p}}{M_{\rm J}}}^{3/2}
            \p{\frac{m}{4 M_{\oplus}}}^{1/2}\nonumber\\
        &\times \p{\frac{m_\star}{M_{\odot}}}^{-3/2}
            \p{\frac{a}{0.4\;\mathrm{AU}}}^6
            \p{\frac{a_{\rm p}}{5\;\mathrm{AU}}}^{-9/2}
            \p{\frac{R}{2 R_{\oplus}}}^{-3/2}.
\end{align}
Since $t_{\rm s} \gg t_{\rm s, c}$, we see that tCE2 is stable. Thus, we predict
that a significant fraction of SEs with exterior CJ companions can have
long-lived, significant obliquities ($\gtrsim 60^\circ$) due to being trapped in
tCE2.

% \textcolor{red}{Just a brief investigation:} Finally, we briefly consider the
% possibility of tidal heating. Coarsely, we have that
% \begin{align}
%     \frac{\dot{a}}{a} &\sim -\frac{1}{10^{14}\;\mathrm{yr}}\p{1 -
%         \frac{\Omega_{\rm s}}{n}\cos \theta},\\
%     \dot{E}_{\rm orb} = \frac{GM_{\star}m}{2a}\frac{\dot{a}}{a}
%         &\approx 10^{20}\;\mathrm{erg/s}\p{1 -
%         \frac{\Omega_{\rm s}}{n}\cos \theta}.
% \end{align}
% The black body luminosity of the planet is expected to be $4 \times
% 10^{25}\;\mathrm{erg/s}$ for the fiducial parameters, and so can probably be
% ignored.

\subsection{Application to Ultra-short-period Planet Formation
}\label{ss:disc_usp}

In \citet{millholland2020formation}, the authors consider capture into the CS2
resonance as part of a mechanism to form ultra-short period planets (USPs). They
envision a process consisting of three stages:
\begin{itemize}
    \item Rapid initial guaranteed capture into CS2.

    \item Runaway inward tidal migration due to the simultaneous increase of the
        CS2 obliquity during migration.

    \item Stalling of inward migration when tidal dissipation becomes
        sufficiently strong to destroy CS2.
\end{itemize}
They consider an inner proto-USP with two external perturbing planets, and
imagine that the spin of the planet quickly evolves to its equilibrium value,
i.e.\ satisfying Eq.~\eqref{eq:weaktide_dWszero}. In our language, this
corresponds to immediate evolution into tCE2. In this section, we revisit this
mechanism with the results presented earlier in this paper.

First, a proto-USP evolves from its initial condition into one of the tCE on the
timescale
% 1 / (1.5 * 1e-3 * (Msun / (8 Mearth)) * ((2 Rearth) / (0.035 AU))^3 * 2 * pi / (0.035)^(3/2))
% = 1157
\begin{align}
    \frac{1}{t_{\rm s}} ={}& \frac{1}{1200\;\mathrm{yr}}
            \p{\frac{1}{4k}}
            \p{\frac{2k_2/Q}{10^{-3}}}
            \p{\frac{M_\star}{M_{\odot}}}^{3/2}
            \p{\frac{m}{8 M_{\oplus}}}^{-1}\nonumber\\
        &\times \p{\frac{R}{2R_{\oplus}}}^3
            \p{\frac{a}{0.035\;\mathrm{AU}}}^{-9/2}.
\end{align}
This is much shorter than the age of the system, and so the proto-USP can be
treated as starting in one of the tCE\@.

Next, we investigate the outcome of the initial, rapid spin evolution. First,
due to the small semi-major axis ratio, we must generalize Eq.~\eqref{eq:wlp}
\citep{lai_2017}:
\begin{equation}
    \omega_{\rm lp} = \frac{3m_{\rm p}}{4M_{\star}}
        \alpha^3 n f(\alpha),
        \label{eq:def_g_usp}
\end{equation}
where $\alpha = a / a_{\rm p}$ and \citep{murray1999solar}
\begin{equation}
    f(\alpha) \equiv \frac{b_{\rm 3/2}^{(1)}(\alpha)}{3\alpha}
        \approx 1 + \frac{15}{8}\alpha^2 + \frac{175}{64}\alpha^4 + \dots,
\end{equation}
where $b_{\rm 3/2}^{(1)}$ is a Laplace coefficient. With this modification,
$\eta_{\rm sync}$ (Eq.~\ref{eq:def_etasync}) can be computed:
% (1 + 15/8 * (1.3)^(-4/3) + 175/64 * 1.3^(-8/3)) * 0.012
\begin{align}
    \eta_{\rm sync} ={}& 0.027 \cos I
            \p{\frac{k}{k_{\rm q}}}
            \p{\frac{m m_{\rm p}}{\p{8 M_{\oplus}}^2}}
            \p{\frac{M_\star}{M_{\odot}}}^{-2}
            \p{\frac{a}{0.035 \;\mathrm{AU}}}^{3}\nonumber\\
        &\times \p{\frac{\alpha^3 f(\alpha)}{0.016}}
            \p{\frac{R}{2R_{\oplus}}}^{-3}.
            \label{eq:etasync_usp}
\end{align}
We have evaluated using an $\alpha$ corresponding to a period ratio $P_{\rm p} /
P = 1.5$. For $I = 5^\circ$, roughly consistent with typical proto-USP systems
\citep{dai2018larger}, Eq.~\eqref{eq:def_etac} shows that $\eta_{\rm sync} \ll
\eta_{\rm c}$, and thus there are four CSs at spin-orbit synchronization. If,
following \citet{millholland2020formation}, we take the initial planetary
obliquity to be zero, the planet is guaranteed to evolve into tCE1, and not tCE2
(see Fig.~\ref{fig:Hhists_0_06}). If we assume instead a randomly oriented
initial planetary spin, perhaps again due to early giant impacts,
Fig.~\ref{fig:probs5} suggests that the probability of capture into tCE2 is
still only modest ($\lesssim 10\%$). A more sophisticated calculation including
the effect of a third planet does not greatly modify these results. Note that
this calculation is in tension with Fig.~4 of \citet{millholland2020formation},
where they find that $\eta \sim 1$ for the same physical parameters when
$\Omega_{\rm s}$ is at its equilibrium value (i.e.\ satisfying
Eq.~\ref{eq:weaktide_dWszero}, or $\Omega_{\rm s} \approx n$ when the obliquity
is small).

The second stage of the proposed mechanism, runaway inward migration after
attaining tCE2, requires that the initial orbital decay time scale be
sufficiently fast. We find that:
% 1 / (1.5 * 1e-3 * (Msun / (8 Mearth)) * ((2 Rearth) / (0.035 AU))^5 * 2 * pi / (0.035)^(3/2))
% = 1.955e8
\begin{align}
    -\frac{\dot{a}}{a} ={}& \frac{1}{2 \times 10^8\;\mathrm{yr}}
            \p{1 - \frac{\Omega_{\rm s}}{n}\cos \theta}
            \p{\frac{2k_2/Q}{10^{-3}}}
            \p{\frac{M_\star}{M_{\odot}}}^{3/2}\nonumber\\
        &\times \p{\frac{m}{8M_{\oplus}}}^{-1}
            \p{\frac{R}{2R_{\oplus}}}^5
            \p{\frac{a}{0.035\;\mathrm{AU}}}^{-13/2}.
            \label{eq:adot_usp}
\end{align}
For $\eta_{\rm sync} \ll \eta_{\rm c}$, we can show that $\p{1 - \Omega_{\rm
s}\cos \theta / n} \approx 1$ in tCE2 by using Eqs.~\eqref{eq:def_tce2_approx},
so indeed the orbit of the proto-USP is able to decay within the lifetime of the
system. On the other hand, in tCE1, $\omega_{\rm s} \approx n$ and $\cos \theta
\approx 1 - \eta_{\rm sync}^2 \sin^2 I / 2$, so $\dot{a} / a$ is suppressed by a
factor of $\sim \eta_{\rm sync}^2 \sin^2 I$. This shows that a proto-USP in tCE1
is unable to initiate runaway orbital decay within the age of the system.
Note that this constraint also implies $\eta_{\rm sync}$
(Eq.~\ref{eq:etasync_usp}) cannot be increased by considering proto-USPs with
larger values of $a$, as the initial orbital decay will become too slow.

Finally, we compute the orbital separation at which tCE2 becomes unstable due to
sufficiently strong tidal dissipation. Evaluating Eq.~\eqref{eq:def_ts_crit}, we
find that tCE2 breaks ($t_{\rm s} \lesssim t_{\rm s, c}$) when the semi-major
axis is smaller than $a_{\rm break}$, where
% (1e-3 * (Msun / (8 Mearth))^3 * ((2 Rearth) * (0.05 AU))^(9/2))^(1/9) / (1 AU)
% = 0.033198
\begin{align}
    a_{\rm break} \approx{}& 0.033 \;\mathrm{AU}
        \p{\frac{2k_2/Q}{10^{3}}}^{-1/9}
        \p{\frac{M_\star}{M_{\odot}}}^{1/3}
        \p{\frac{mm_{\rm p}}{\p{8M_{\oplus}}^2}}^{-1/6}\nonumber\\
        &\times \p{\frac{R}{2R_{\oplus}}}^{1/2}
        \p{\frac{a_{\rm p}}{0.05\;\mathrm{AU}}}^{1/2}
        \p{\frac{1}{\sin I \cos^2 I}}^{1/9}.
\end{align}
This corresponds to a 2.7-day orbit. Once the system exits tCE2, it rapidly
evolves to tCE1, which suppresses its orbital decay (Eq.~\ref{eq:adot_usp}).
This final orbital separation does not qualify as a USP ($P \lesssim
\mathrm{day}$).

In summary, our results for the formation of USPs via runaway tidal migration
are in tension with those of \citet{millholland2020formation}. In particular, we
find that only proto-USPs with a large primordial obliquity have a probability
of evolving into tCE2 initially. Furthermore, proto-USPs that successfully
initiate runaway tidal migration after reaching tCE2 will cease their inward
migration before becoming a USP\@.

\subsection{Application to WASP-12b}\label{ss:disc_wasp12b}

In \citet{millholland2019obliquity}, the authors consider the possibility that
the measured orbital decay of the HJ WASP-12b \citep[$P / \dot{P} = -3.2
\;\mathrm{Myr}$; ][]{maciejewski2016departure, patra2017apparently} is caused by
tidal dissipation in the HJ enhanced by being trapped in a high-obliquity CS due
to an undetected exterior planet. In order for this scenario to be achieved,
many conditions are necessary:
\begin{itemize}
    \item The HJ must have had a sufficiently small initial semimajor axis such
        that its orbital decay timescale is less than the age of the system.

    \item The exterior planet must be sufficiently massive to keep the HJ in a
        high-obliquity tCE today.

    \item The RV signal of the exterior planet must be smaller than the
        residuals of the published RVs, $\sim 16\;\mathrm{m/s}$
        \citep{hebb2009wasp, husnoo2011orbital, knutson2014friends,
        bonomo2017gaps}.

    \item The initial configuration of the HJ and exterior planet must be
        dynamically stable.
\end{itemize}
Finally, if all four of these conditions are satisfied, the $\eta_{\rm sync}$ of
the initial configuration can be used to estimate the likelihood of the spin of
the HJ initally evolving into tCE2. We use the following current-day system
parameters \citet{hebb2009wasp, maciejewski2013multi, millholland_wasp12b}:
$R_\star = 1.63R_{\odot}$, $M_\star = 1.36M_{\odot}$, $R = 1.89R_{\rm J}$,
$M_{\rm p} = 1.41M_{\rm J}$, and $a = 0.023 \;\mathrm{AU}$.

First, we require that the initial HJ semimajor axis, denoted $a_{\rm i}$, is
sufficiently small that decay occurs within the age of the system. The semimajor
axis decay rate can be expressed as
% 1 / (1.5 * 1e-7 * (1.36 Msun) / (1.41 Mjup) * ((1.89 Rjup) / (0.038 AU))^5 * (G * (1.36 Msun) / (0.038 AU)^3)^(1/2) * (1 yr))
% = 9.82e8
\begin{align}
    \frac{\dot{a}}{a_{\rm i}}
        =& -\frac{1}{\mathrm{Gyr}}
            \p{\frac{2k_2/Q}{10^{-6}}}
            \p{\frac{M_\star}{1.36 M_{\odot}}}^{3/2}
            \p{\frac{m}{1.41M_J}}^{-1}\nonumber\\
        &\times \p{\frac{R}{1.89R_j}}^5
            \p{\frac{a_{\rm i}}{0.038\;\mathrm{AU}}}^{-13/2}
            \p{1 - \frac{\Omega_{\rm s}}{n}\cos \theta}.
            \label{eq:hj_adot}
\end{align}
Again, $t_{\rm s} \ll a_{\rm i} / \dot{a}$, so we can treat the HJ as
immediately evolving into a tCE\@. Thus, we see that the initial semimajor axis
for the HJ cannot exceed $0.038\;\mathrm{AU}$ even when $\p{1 - \Omega_{\rm
s}\cos \theta/n} \approx 1$, as is the case for rapid evolution into tCE2 while
$\eta_{\rm sync} \ll 1$.

Next, we address the constraint that tCE2 is stable under the influence of the
exterior planet. Due to the fact that the angular momentum of the HJ greatly
exceeds that of the exterior planet ($L \gg L_{\rm p}$; we will verify this
retroactively), Eq.~\eqref{eq:dldt1} needs to be modified. The precession of
$\uv{l}$ about $\bm{J} = J\uv{j} \equiv \bm{L} + \bm{L}_{\rm p}$, the total
angular momentum axis, can be written as
\begin{equation}
    \rd{\uv{l}}{t} = \omega_{\rm lp}\cos I \frac{J}{L_{\rm p}}
        \p{\uv{l} \times \uv{j}},
\end{equation}
where $\omega_{\rm lp}$ is still given by Eq.~\eqref{eq:wlp}. Thus, we see that
the precession frequency $\abs{g}$ is enhanced by a factor of $J / L_{\rm p}
\approx L / L_{\rm p}$. With this modification, the stability of tCE2 requires
\begin{equation}
    \frac{1}{t_{\rm s}} \lesssim \omega_{\rm lp}\cos I \frac{L}{L_{\rm p}}
        \sin I \sqrt{\frac{\eta_{\rm sync}\cos I}{2}}.
\end{equation}
This then yields that
\begin{align}
    \frac{a_{\rm p}}{a}
        \lesssim{}& 3.3 \p{\sin I \cos^2 I}^{4/21}
            \p{\frac{m}{1.41 M_J}}^{12/21}
            \p{\frac{M_\star}{1.36 M_{\odot}}}^{-12/21}\nonumber\\
        &\times \p{\frac{2k_2/Q}{10^{-6}}}^{-4/21}
            \p{\frac{a}{0.023\;\mathrm{AU}}}^{6/7}
            \p{\frac{R}{1.89 R_J}}^{-6/7}.\label{eq:hj_tce2_stab}
\end{align}
Notably, this expression does not depend on the perturber mass but only on
$a_{\rm p}$.

Next, we address the constraint that the RV signal of the exterior planet be
smaller than $16\;\mathrm{m/s}$, which requires
% (G * (1.36 Msun) / (0.125 AU))^(1/2) * (105 Mearth) / (1.36 Msun) / 2^(1/2)
% 16.106 m/s
\begin{equation}
    \p{\frac{a_{\rm p}}{0.125\;\mathrm{AU}}}^{-1/2}
    \p{\frac{m_{\rm p}}{105M_{\oplus}}}
    \p{\frac{M_\star}{1.36M_{\odot}}}^{-1/2}
    \p{\frac{\sin i}{1 / \sqrt{2}}} \lesssim 1,
    \label{eq:hj_rv}
\end{equation}
where $i$ is the line-of-sight inclination angle. Here, we have taken $a_{\rm
p}$ to be the maximum value permitted by the combination of
Eqs.~(\ref{eq:hj_adot},~\ref{eq:hj_tce2_stab}). For these values of $a_{\rm p}$
and $m_{\rm p}$, we still have $L / L_{\rm p} \sim 0.4$, and so $L \gg L_{\rm
p}$ is be well satisfied for the permitted parameter space.

Finally, we require that the initial orbital configuration be dynamically
stable. We use the Hill stability criterion \citep[e.g.][]{petit2020path}:
\begin{equation}
    \frac{a_{\rm p} - a_{\rm i}}{a_{\rm i}} > 2\sqrt{3}\frac{a_{\rm p} + a_{\rm
        i}}{2a_{\rm i}} \p{\frac{m + m_{\rm p}}{3M_{\star}}}^{1/3}.
\end{equation}
Assuming $m_{\rm p} \ll m$, this yields:
\begin{equation}
    \frac{a_{\rm p}}{a_{\rm i}} > 1.29.\label{eq:hj_orbstab}
\end{equation}

By using the constraints given by Eqs.~(\ref{eq:hj_adot},~\ref{eq:hj_orbstab}),
the initial $\eta_{\rm sync, i}$ can be constrained:
\begin{align}
    \eta_{\rm sync, i} <{}& 0.016 \cos I\p{\frac{m}{1.41M_J}}^2
            \p{\frac{M_\star}{1.36M_{\odot}}}^{-2}
            \p{\frac{a_{\rm i}}{0.038\;\mathrm{AU}}}^{3}\nonumber\\
        &\times \p{\frac{a_{\rm p} / a_{\rm i}}{1.29}}^{-7/2}
            \p{\frac{R}{1.89R_{\rm J}}}^{-3}.\label{eq:hj_etasync}
\end{align}
It is possible that $R$ is larger today than its primordial value due to
inflation induced by increased stellar irradiation. However, if a smaller value
of $R$ is used in Eq.~\eqref{eq:hj_etasync}, the value of $a_{\rm i}$ must also
be decreased such that $R^5 / a_{\rm i}^{13/2}$ is constant
(Eq.~\ref{eq:hj_adot}), which further decreases $\eta_{\rm sync, i}$.

For $\eta_{\rm sync} = 0.016$, we find that there are four CSs unless $I$ is
extremely small. From Fig.~\ref{fig:Hhists_0_06}, we can infer that prograde
primordial obliquities will evolve towards tCE1 in the regime where $\eta_{\rm
sync} \ll \eta_{\rm c}$. On the other hand, if the primordial obliquity of the
HJ is assumed to be isotropically distributed, then Fig.~\ref{fig:probs20}
suggests that the probability of entry into tCE2 is still $\lesssim 2\%$ even if
the perturbing planet is misaligned by $I = 20^\circ$.

In summary, the joint constraints we have placed on the WASP-12b \& perturber
system imply that the initial $\eta_{\rm sync, i}$ is unlikely to have been
large. This suggests that early capture into tCE2 is unlikely from either an
isotropic or prograde-favoring initial obliquity distribution. However, it is
possible that additional mechanisms can generate a favorable initial obliquity
distribution, such as the presence of a dissipating protoplanetary disk in the
early lifetime of the system quickly driving the obliquity of WASP-12b to $\sim
90^\circ$ \citep{millholland_disk, su2020}.

\section{Summary and Discussion}\label{s:summary}

In this work, we have studied the evolution of an interior planet's spin due to
the joint effect of tidal dissipation and gravitational interactions with an
exterior perturber. Our results extend our study of Colombo's Top in a previous
paper \citep{su2020} to include a dissipative torque, for which we have adopted
the weak friction theory of the equilibrium tide \citep{lai2012}. Our primary
esults are as follows:
\begin{itemize}
    \item We calculate the modification of the equilibria of Colombo's Top
        (Cassini States; CSs) due to tidal dissipation and evaluate the
        stability of the CSs in the presence of the tidal dissipation.

    \item We show that, for a given orbital architecture, the spin of the inner
        planet has at most two spin configurations (spin frequency and
        orientation) that are stable under the effect of tidal dissipation. We
        call these equilibria \emph{tidal Cassini Equilibria} (tCE\@; see
        Fig.~\ref{fig:6equils}). The locations of these equilibria are
        determined by the system architecture and are parameterized by
        $\eta_{\rm sync}$ (Eq.~\ref{eq:def_etasync}).

    \item We show that which tCE a given initial planetary spin configuration
        reaches can be probabilistic for certain initial conditions. For these
        initial conditions, the probabilities of approaching either tCE can be
        determined by careful study of the dynamics upon separatrix encounter
        (Sections~\ref{ss:toy_outcomes} and~\ref{ss:phop_weaktide}). As an
        example, if the initial spin orientation is assumed to be isotropically
        distributed, the probabilities of entering tCE1 and tCE2 are given in
        Figs.~\ref{fig:probs20} and~\ref{fig:probs5} as a function of $\eta_{\rm
        sync}$ for two different mutual inclinations $I$.

    \item In three exoplanetary scenarios of interest, application of our
        results yields the following conclusions: (i) a super-Earth with an
        exterior cold Jupiter has substantial probability of being trapped in a
        high-obliquity tCE\@; (ii) formation of ultra-short-period planets via
        tidal runaway is unlikely due to low probability of capture into a
        high-obliquity tCE\@; and (iii) WASP-12b is unlikely to be undergoing
        enhanced orbital decay due to obliquity tides, as the probability of
        entry into tCE2 is also low.
\end{itemize}

Above, we have studied tidal dissipation in the planet via the weak friction
theory of the equilibrium tide \citep{lai2012}. However, other mechanisms of
tidal dissipation may be dominant, depending on the system parameters
\citep[e.g.][]{papaloizou_ivanov_inertial, teyssandier2019formation}. While the
detailed outcomes may change, a different tidal mechanism is amenable to the
same analysis presented in this paper: The tCE can still be found by an analysis
mirroring that shown in Fig.~\ref{fig:6equils}, and the probabilistic outcome of
a separatrix encounter can still be solved using the techniques of
Sections~\ref{ss:toy_outcomes} and~\ref{ss:phop_weaktide}.

Additionally, we have primarily focused on an isotropic distribution of initial
spin orientation, assuming that giant impacts will effectively randomize a
planet's spin. More physically accurate distributions can be used in the case of
planetary mergers \citep{li2020planetary} or many smaller impacts
\citep{dones1993does}. Figures~\ref{fig:probs20}--\ref{fig:probs5} can be
updated accordingly by convolving the distributions in the right panels of
Figs.~\ref{fig:Hhists_0_06}--\ref{fig:Hhists_0_70} with any initial obliquity
distribution. The qualitative results are unlikely to change, while the detailed
probabilities for tCE2 capture can increase [decrease] if the initial obliquity
distribution favors [disfavors] $\theta_{\rm i} \approx 90^\circ$ compared to
the isotropic distribution.

\section{Acknowledgements}

We thank Sarah Millholland for useful discussions. This work has been supported
in part by NSF grant AST1715246. YS is supported by the NASA FINESST grant
19-ASTRO19-0041.%chktex 8

\bibliography{Su_weak_tides}
\bibliographystyle{aasjournal}

\appendix

\onecolumn

\section{Convergence of Initial Conditions Inside the Separatrix to CS2
}\label{app:cs_stab2}

In Section~\ref{ss:tidal_equils}, we studied the stability of the CSs under of
tidal alignment torque given by Eq.~\eqref{eq:dsdt_tide_toy}, finding that CS2
is locally stable. Later, in Section~\ref{ss:toy_outcomes}, we found that all
initial conditions within the separatrix converge to CS2, which is not
guaranteed by local stability of CS2. In this section, we give an analytic
demonstration that all points inside the separatrix indeed converge to CS2,
focusing on the case where $\eta \ll 1$.

Similarly to the analytic calculation in Section~\ref{ss:toy_outcomes}, we seek
to compute the change in the unperturbed Hamiltonian over a single libration
cycle. To calculate the evolution of $H$, we first parameterize the unperturbed
trajectory (similarly to Eq.~\ref{eq:sep_theta}). For initial conditions inside
the separatrix, the value of $H$ can be written $H = H_{\rm sep} + \Delta H$
where $\Delta H > 0$, and the two legs of the libration trajectory can be
written:
\begin{align}
    \cos \theta_{\pm} &\approx
        \eta \cos I \pm \sqrt{2\eta\s{\sin I\p{1 - \cos \phi} - \Delta H}}.
        \label{eq:lib_cycle_toy}
\end{align}
We have taken $\sin \theta \approx 1$, a good approximation in zone II when
$\eta \ll 1$. Note that there are some values of $\phi$ for which no solutions
of $\theta$ exist, reflecting the fact that the libration cycle does not extend
over the full interval $\phi \in [0, 2\pi]$. During a libration cycle,
$\theta_-$ [$\theta_+$] is traversed while $\phi' > 0$ [$\phi' < 0$], i.e.\ the
trajectory librates counterclockwise in $(\cos \theta, \phi)$ phase space (see
Fig.~\ref{fig:1contours}).

The leading order change to $H$ over a single libration cycle can then computed by
integrating $\rdil{H}{t}$ along this trajectory, yielding:
\begin{align}
    \oint \rd{H}{t}\;\mathrm{d}t
        &= \oint \p{\rd{(\cos \theta)}{t}}_{\rm tide}
            \;\mathrm{d}\phi,\nonumber\\
        &= \int\limits_{\phi_{\min}}^{\phi_{\max}}
                \frac{1}{t_{\rm s}}
                \p{\sin^2\theta_- - \sin^2\theta_+} \;\mathrm{d}\phi\nonumber\\
        &= \frac{1}{t_{\rm s}}
            \int\limits_{\phi_{\min}}^{\phi_{\max}}
                4\eta \cos I \sqrt{2\eta\s{\sin I\p{1 - \cos \phi} - \Delta H}}
                \;\mathrm{d}\phi > 0.
\end{align}
Here, $\phi_{\min} > 0$ and $\phi_{\max} < 2\pi$ are defined such that the
trajectory librates over $\phi \in \s{\phi_{\min}, \phi_{\max}}$. Thus, $H$ is
strictly increasing for all initial conditions inside the separatrix, and they
all converge to CS2.

% \section{Separatrix Crossing Probability with Weak Tidal Friction
% }\label{app:sep_crossing_dynamics}

% In this appendix, we analytically calculate $\Delta K_{\pm}$ for use in
% Eq.~\eqref{eq:def_pc_weaktide} to calculate the probabilities of the two
% possible transitions upon separatrix encounter. We first rewrite the full
% equations of motion for the planet's spin including weak tidal friction in
% component form:
% \begin{align}
%     \rd{\theta}{t} &= g\sin I \sin \phi -
%         \frac{1}{t_{\rm a}} \sin \theta,\label{eq:ds_fullq}\\
%     \rd{\phi}{t} &= -\alpha\cos\theta
%         - g\p{\cos I + \sin I \cot \theta \cos \phi}\label{eq:ds_fullphi},\\
%     \rd{\Omega_{\rm s}}{t}
%         &= \frac{1}{t_{\rm a}} \s{2n \cos \theta
%             - g\Omega_{\rm s}\p{1 + \cos^2\theta}}\label{eq:ds_fulls}.
% \end{align}

% Note that Eq.~\eqref{eq:def_dK_weaktide} can be rewritten
% \begin{align}
%     \Delta K_{\pm} &= \oint_{\mathcal{C}_{\pm}} \rd{H}{t}
%         - \rd{H_{\rm sep}}{t}\;\mathrm{d}t
%         \nonumber\\
%         &= \oint_{\mathcal{C}_{\pm}}
%            \p{\rd{(\cos\theta)}{t}}_{\rm tide}
%             + \frac{\dot{\Omega}_{\rm s}}{\dot{\phi}}
%             \p{\pd{H}{s} - \pd{H_{\rm sep}}{s}}\;\mathrm{d}\phi
%                 \label{eq:app_dhpm}.
% \end{align}

% We then evaluate $\Delta K_{\pm}$ by integrating along the two legs of the
% separatrix $\mathcal{C}_{\pm}$ (see Fig.~\ref{fig:1contours}). Note that we must
% use the value of $\eta$ at the moment of separatrix encounter, as the evolution
% of $\Omega_{\rm s}$ changes the spin-orbit precession frequency $\alpha$
% \begin{align}
%     \Delta K_{\pm} ={}& \oint_{\mathcal{C}_{\pm}}
%         \frac{1}{t_{\rm s}}\sin^2\theta
%             \p{\frac{2n}{\Omega_{\rm s}} - \cos\theta}
%          +
%             \s{\frac{\alpha}{\Omega_{\rm s}}
%                 \frac{\cos^2\theta}{2} + g\frac{\Omega_{\rm
%             c}}{2\Omega_{\rm s}^2}\cos^2 I}\rd{\Omega_{\rm
%             s}}{t}\rd{t}{\phi}\;\mathrm{d}\phi\\
%      \approx{}& \oint_{\mathcal{C}_{\pm}}
%         \frac{1}{t_{\rm s}}\sin^2\theta
%             \p{\frac{2n}{\Omega_{\rm s}} - \cos\theta}
%         + \frac{1}{t_{\rm s}}\frac{n}{\Omega_{\rm s}\eta}
%             \s{\p{\cos \theta}_{\mathcal{C}_{\pm}} - \frac{\Omega_{\rm s}}{2n}}
%             \s{2 \cos I \pm \sqrt{2 \sin I\p{1 - \cos \phi} / \eta}}
%             \;\mathrm{d}\phi.\label{eq:app_deltaK}
%     t_{\rm s} \Delta K_{\pm} \approx{}&
%         -2\cos I\p{\pm 2\pi \eta \cos I + 8\sqrt{\eta \sin I}}
%         \pm 2\pi \frac{\Omega_{\rm s}}{n}\cos I
%         - 8\eta \cos I \sqrt{\sin I / \eta}
%             + \frac{4\Omega_{\rm s}}{n}\sqrt{\sin I/\eta}\nonumber\\
%         &+ \frac{2n}{\Omega_{\rm s}}\p{\mp 2\pi\p{1 - 2\eta \sin I}
%             + 16\cos I \eta^{3/2}\sqrt{\sin I}}
%             + 8\sqrt{\eta \sin I}
%             \pm 2 \pi \eta \cos I
%             - \frac{64}{3} \p{\eta \sin I}^{3/2}.\label{eq:app_deltaK}
% \end{align}
% At this point, since $\eta$ often is not small, it is necessary to integrate
% along the numerically computed $\mathcal{C}_{\pm}$ (instead of the approximate
% solution given by Eq.~\ref{eq:sep_theta}) to obtain sufficiently accurate values
% for $\Delta K_{\pm}$. These numerically integrated values are used in
% Eq.~\eqref{eq:def_pc_weaktide} to semi-analytically compute the tCE2
% probabilities (blue curves) in
% Figs.~\ref{fig:pc_fits_0_06}--\ref{fig:pc_fits_0_20}.

\label{lastpage} % chktex 24
\end{document}